---
title: "Week 2 - Homework"
author: "STAT 420, Summer 2018, Connor Segneri - segneri3"
date: ''
output:
  pdf_document: default
urlcolor: cyan
---


***

## Exercise 1 (Using `lm`)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

```{r message=FALSE, warning=FALSE}
library(MASS)
```

**(a)** Suppose we would like to understand the size of a cat's heart based on the body weight of a cat. Fit a simple linear model in `R` that accomplishes this task. Store the results in a variable called `cat_model`. Output the result of calling `summary()` on `cat_model`.

```{r message=FALSE, warning=FALSE}
cat_model = lm(Hwt ~ Bwt, data=cats)
summary(cat_model)
```

**(b)** Output only the estimated regression coefficients. Interpret $\hat{\beta_0}$ and $\beta_1$ in the *context of the problem*. Be aware that only one of those is an estimate.

```{r message=FALSE, warning=FALSE}
cat_model$coefficients
```

$\hat{\beta_0}$, **-0.3566624** is supposed to represent the heart weight (g) of a cat when the weight (kg) is zero. Obviously the height and weight of a cat must be greater than zero, so any predictions made can be disregarded until they are made within the estimated range of the cat's body weight: **```r range(cats$Bwt)```**.

$\beta_1$, 4.0340627, represents the relationship between the cat's height and body weight. Based on the data, the linear model estimates that with 1 kg in body weight gained by a cat, the expected heart weight of the cat increases by 1 g.

**(c)** Use your model to predict the heart weight of a cat that weights **2.7** kg. Do you feel confident in this prediction? Briefly explain.

```{r message=FALSE, warning=FALSE}
predict(cat_model, newdata=data.frame(Bwt = 2.7))
```

I feel confident about this prediction, as it was made within the range of the estimation data. The range of the estimation data, the body weight (kg) of the cat, is **```r range(cats$Bwt)```**.

**(d)** Use your model to predict the heart weight of a cat that weights **4.4** kg. Do you feel confident in this prediction? Briefly explain.

```{r message=FALSE, warning=FALSE}
predict(cat_model, newdata=data.frame(Bwt = 4.4))
```

I do not feel as condfident in this prediction. The relationship between the independent and dependent variables can change outside of the estimation range, which we previously established to be **```r range(cats$Bwt)```** for a cat's body weight (kg).

**(e)** Create a scatterplot of the data and add the fitted regression line. Make sure your plot is well labeled and is somewhat visually appealing.

```{r message=FALSE, warning=FALSE, fig.align='center'}
library(ggplot2)

ggplot(data=cats, aes(x=Bwt,y=Hwt)) +
  geom_point(color = 'red') +
  labs(x = "Cat's Body Weight (kg)", y = "Cat's Heart Weight (g)") +
  geom_smooth(method="lm", col="blue")
```

**(f)** Report the value of $R^2$ for the model. Do so directly. Do not simply copy and paste the value from the full output in the console after running `summary()` in part **(a)**.

```{r warning=FALSE, message=FALSE}
summary(cat_model)$r.squared
```

***

## Exercise 2 (Writing Functions)

This exercise is a continuation of Exercise 1.

**(a)** Write a function called `get_sd_est` that calculates an estimate of $\sigma$ in one of two ways depending on input to the function. The function should take three arguments as input:

- `fitted_vals` - A vector of fitted values from a model
- `actual_vals` - A vector of the true values of the response
- `mle` - A logical (`TRUE` / `FALSE`) variable which defaults to `FALSE`

The function should return a single value:

- $s_e$ if `mle` is set to `FALSE`.
- $\hat{\sigma}$ if `mle` is set to `TRUE`.

```{r message=FALSE, warning=FALSE}
get_sd_est = function(fitted_vals, actual_vals, mle=FALSE) {
  if (mle){
    sqrt((1/length(actual_vals)) * sum((actual_vals - fitted_vals)^2))
  } else {
    sqrt((1/(length(actual_vals)-2)) * sum((actual_vals - fitted_vals)^2))
  }
}
```

**(b)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `FALSE`. Explain the resulting estimate in the context of the model.

```{r message=FALSE, warning=FALSE}
actual_values = cats$Hwt
fitted_values = predict(cat_model, newdata = cats)

get_sd_est(fitted_vals = fitted_values, actual_vals = actual_values)
```

This means that $s_e$, or the standard error of estimate is **1.452373**. This measures the variation in the actual values of the cat's heart weight to the computed values of the cat's heart weight using the linear model **cat_model**. The smaller the value is the better, if the standard error of estimate is zero, then the correlation is perfect.

**(c)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `TRUE`. Explain the resulting estimate in the context of the model. Note that we are trying to estimate the same parameter as in part **(b)**.

```{r message=FALSE, warning=FALSE}
get_sd_est(fitted_vals = fitted_values, actual_vals = actual_values, mle = TRUE)
```

The estimated standard deviation, or $\hat{\sigma}$, is **1.442252**. The difference between this value and the standard error of estimate is minor. However, standard error of estimation is considered unbiased, since $E[s_e]=\sigma^2$, and the estimated standard deviation is considered biased because $\hat{\sigma}\neq\sigma^2$.

**(d)** To check your work, output `summary(cat_model)$sigma`. It should match at least one of **(b)** or **(c)**.

```{r message=FALSE, warning=FALSE}
summary(cat_model)$sigma
```

As expected, $E[s_e]=\sigma^2$.

***

## Exercise 3 (Simulating SLR)

Consider the model

\[
Y_i = 5 + -3 x_i + \epsilon_i
\]

with 

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 10.24)
\]

where $\beta_0 = 5$ and $\beta_1 = -3$.

This exercise relies heavily on generating random observations. To make this reproducible we will set a seed for the randomization. Alter the following code to make `birthday` store your birthday in the format: `yyyymmdd`. For example, [William Gosset](https://en.wikipedia.org/wiki/William_Sealy_Gosset), better known as *Student*, was born on June 13, 1876, so he would use:

```{r}
birthday = 19951015
set.seed(birthday)
```

**(a)** Use `R` to simulate `n = 25` observations from the above model. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 25, 0, 10)
```

You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Store the data frame this function returns in a variable of your choice. Note that this function calls $y$ `response` and $x$ `predictor`.

```{r message=FALSE, warning=FALSE}
n_obs = 25
beta_0 = 5
beta_1 = -3
sigma = sqrt(10.24)

sim_slr = function(x, beta_0 = 10, beta_1 = 5, sigma = 1) {
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, response = y)
}

xy_predicted_pairs = sim_slr(x, beta_0=beta_0, beta_1=beta_1, sigma=sigma)
xy_predicted_pairs
```

**(b)** Fit a model to your simulated data. Report the estimated coefficients. Are they close to what you would expect? Briefly explain.

```{r message=FALSE, warning=FALSE}
model = lm(response ~ predictor, data=xy_predicted_pairs)

summary(model)
```

This is about what I expected, the only strange thing is how far off the y-intercept is from the original model. Original was `5` and the model gives `6.9846` for the intercept. The slope is only `0.3` off.

**(c)** Plot the data you simulated in part **(a)**. Add the regression line from part **(b)** as well as the line for the true model. Hint: Keep all plotting commands in the same chunk.

```{r message=FALSE, warning=FALSE, fig.align='center'}
plot(response ~ predictor, data=xy_predicted_pairs,
     xlab = "Simulated Predictors",
     ylab = "Response")
abline(model, col="red")
abline(beta_0, beta_1, col="blue")
legend("topright", c("Estimate", "Truth"), col = c("red", "blue"), lty = c(1,2), lwd = 2)
```

**(d)** Use `R` to repeat the process of simulating `n = 25` observations from the above model $1500$ times. Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. Some hints:

- Consider a `for` loop.
- Create `beta_hat_1` before writing the `for` loop. Make it a vector of length $1500$ where each element is `0`.
- Inside the body of the `for` loop, simulate new $y$ data each time. Use a variable to temporarily store this data together with the known $x$ data as a data frame.
- After simulating the data, use `lm()` to fit a regression. Use a variable to temporarily store this output.
- Use the `coef()` function and `[]` to extract the correct estimated coefficient.
- Use `beta_hat_1[i]` to store in elements of `beta_hat_1`.
- See the notes on [Distribution of a Sample Mean](http://daviddalpiaz.github.io/appliedstats/introduction-to-r.html#distribution-of-a-sample-mean) for some inspiration.

You can do this differently if you like. Use of these hints is not required.

```{r message=FALSE, warning=FALSE}
beta_hat_1 = c()
for (i in 1:1500){
  xy_predicted_pairs = sim_slr(x, beta_0=beta_0, beta_1=beta_1, sigma=sigma)
  model = lm(response ~ predictor, data=xy_predicted_pairs)
  beta_hat_1[i] = coef(model)[2]
}
```

**(e)** Report the mean and standard deviation of `beta_hat_1`. Do either of these look familiar?

```{r warning=FALSE, message=FALSE}
mean(beta_hat_1)
sd(beta_hat_1)
```

The mean of beta_hat_1 looks a lot like beta_1 for the original model, `-3`.

**(f)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.

```{r, message=FALSE, warning=FALSE, fig.align = 'center'}
hist(beta_hat_1)
```

This histogram is bell-shaped. This indicates that the data is unimodal, meaning that the data has a single mode, identified by the 'peak' of the curve. The histogram looks this way because the data set is normally distributed.

***

## Exercise 4 (Be a Skeptic)

Consider the model

\[
Y_i = 3 + 0 \cdot x_i + \epsilon_i
\]

with

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 4)
\]

where $\beta_0 = 3$ and $\beta_1 = 0$.

Before answering the following parts, set a seed value equal to **your** birthday, as was done in the previous exercise.

```{r}
birthday = 19951015
set.seed(birthday)
```

**(a)** Use `R` to repeat the process of simulating `n = 75` observations from the above model $2500$ times. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 75, 0, 10)
```

Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Hint: Yes $\beta_1 = 0$

```{r message=FALSE, warning=FALSE}
n_obs = 75
beta_0 = 3
beta_1 = 0
sigma = sqrt(4)

beta_hat_1 = c()
for (i in 1:2500){
  xy_predicted_pairs = sim_slr(x, beta_0=beta_0, beta_1=beta_1, sigma=sigma)
  model = lm(response ~ predictor, data=xy_predicted_pairs)
  beta_hat_1[i] = coef(model)[2]
}
```

**(b)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.

```{r, message=FALSE, warning=FALSE, fig.align = 'center'}
hist(beta_hat_1)
```

This histogram is left-skewed. The mean of the data is closer to the right than either the median of the mode.

**(c)** Import the data in [`skeptic.csv`](skeptic.csv) and fit a SLR model. The variable names in `skeptic.csv` follow the same convention as those returned by `sim_slr()`. Extract the fitted coefficient for $\beta_1$.

```{r message=FALSE, warning=FALSE}
library(readr)
skeptics = read_csv('skeptic.csv')

skeptics_model = lm(response ~ predictor, data=skeptics)

beta_1 = coef(skeptics_model)[2]
beta_1
```

**(d)** Re-plot the histogram from **(b)**. Now add a vertical red line at the value of $\hat{\beta_1}$ in part **(c)**. To do so, you'll need to use `abline(v = c, col = "red")` where `c` is your value.

```{r, message=FALSE, warning=FALSE, fig.align = 'center'}
hist(beta_hat_1)
abline(v=beta_1, col="red")
```

**(e)** Your value of $\hat{\beta_1}$ in **(c)** should be negative. What proportion of the `beta_hat_1` values is smaller than your $\hat{\beta_1}$? Return this proportion, as well as this proportion multiplied by `2`.

```{r warning=FALSE, message=FALSE}
prop_smaller = length(beta_hat_1[beta_hat_1 < beta_1]) / length(beta_hat_1)
prop_smaller
prop_smaller * 2
```

**(f)** Based on your histogram and part **(e)**, do you think the [`skeptic.csv`](skeptic.csv) data could have been generated by the model given above? Briefly explain.

No I don't thnk the [`skeptic.csv`](skeptic.csv) data could have been generated by the model given above. The slope of `skeptics model`, $\hat{\beta_1}$, generated from the skeptics data is far from where the bulk of the generated beta_1 values are in `beta_hat_1`. Looking at the histogram tells us that much. An incredibly low proportion of of the `beta_hat_1` values are smaller than $\hat{\beta_1}$. If the skeptics data was generated by the original model, then I would expect the proportion calculated in **e** to be closer to `0.5`. It's still possible that the skeptics data was generated by the original model, but it is improbable.

***

## Exercise 5 (Comparing Models)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will perform some data cleaning before proceeding.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

We have:

- Loaded the data from the package
- Subset the data to relevant variables
    - This is not really necessary (or perhaps a good idea) but it makes the next step easier
- Given variables useful names
- Removed any observation with missing values
    - This should be given much more thought in practice

For this exercise we will define the "Root Mean Square Error" of a model as

\[
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}.
\]

**(a)** Fit three SLR models, each with "ozone" as the response. For the predictor, use "wind speed," "humidity percentage," and "temperature" respectively. For each, calculate $\text{RMSE}$ and $R^2$. Arrange the results in a markdown table, with a row for each model. Suggestion: Create a data frame that stores the results, then investigate the `kable()` function from the `knitr` package.

```{r message=FALSE, warning=FALSE}
library(kableExtra)

wind_model = lm(ozone ~ wind, data=Ozone)
humidity_model = lm(ozone ~ humidity, data=Ozone)
temp_model = lm(ozone ~ temp, data=Ozone)

find_rmse = function(actual, predicted){
  sqrt((1/length(actual)) * sum((actual - predicted)^2))
}

df = data.frame(
  name = c("Wind Model", "Humidity Model", "Temperature Model"),
  rmse = c(
    find_rmse(actual = Ozone$ozone, predicted = predict(wind_model, newdata = Ozone)),
    find_rmse(actual = Ozone$ozone, predicted = predict(humidity_model, newdata = Ozone)),
    find_rmse(actual = Ozone$ozone, predicted = predict(temp_model, newdata = Ozone))
  ),
  r_squared = c(
    summary(wind_model)$r.squared,
    summary(humidity_model)$r.squared,
    summary(temp_model)$r.squared
  )
)

kable(df) %>%
  kable_styling()
```

**(b)** Based on the results, which of the three predictors used is most helpful for predicting ozone readings? Briefly explain.

The predictor that is most helpful for predicting ozone readings is `temperature`. This predictor has the lowest rmse value, and an r_squared value closest to 1.

***

## Exercise 00 (SLR without Intercept)

**This exercise will _not_ be graded and is simply provided for your information. No credit will be given for the completion of this exercise. Give it a try now, and be sure to read the solutions later.**

Sometimes it can be reasonable to assume that $\beta_0$ should be 0. That is, the line should pass through the point $(0, 0)$. For example, if a car is traveling 0 miles per hour, its stopping distance should be 0! (Unlike what we saw in the book.)

We can simply define a model without an intercept,

\[
Y_i = \beta x_i + \epsilon_i.
\]

**(a)** [In the **Least Squares Approach** section of the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#least-squares-approach) you saw the calculus behind the derivation of the regression estimates, and then we performed the calculation for the `cars` dataset using `R`. Here you need to do, but not show, the derivation for the slope only model. You should then use that derivation of $\hat{\beta}$ to write a function that performs the calculation for the estimate you derived. 

In summary, use the method of least squares to derive an estimate for $\beta$ using data points $(x_i, y_i)$ for $i = 1, 2, \ldots n$. Simply put, find the value of $\beta$ to minimize the function

\[
f(\beta)=\sum_{i=1}^{n}(y_{i}-\beta x_{i})^{2}.
\]

Then, write a function `get_beta_no_int` that takes input:

- `x` - A predictor variable
- `y` - A response variable

The function should then output the $\hat{\beta}$ you derived for a given set of data.

**(b)** Write your derivation in your `.Rmd` file using TeX. Or write your derivation by hand, scan or photograph your work, and insert it into the `.Rmd` as an image. See the [RMarkdown documentation](http://rmarkdown.rstudio.com/) for working with images.

**(c)** Test your function on the `cats` data using body weight as `x` and heart weight as `y`. What is the estimate for $\beta$ for this data?

**(d)** Check your work in `R`. The following syntax can be used to fit a model without an intercept:

```{r, eval = FALSE}
lm(response ~ 0 + predictor, data = dataset)
```

Use this to fit a model to the `cat` data without an intercept. Output the coefficient of the fitted model. It should match your answer to **(c)**.

